### 2025-02-22 paper/sections/ExperimentalSetup.tex
add conclusion

\subsection{Data Splits and Usage}
%TODO: refer to dataset used
The dataset was divided into three subsets: training, development (validation), and testing. Specifically, 80\% of the training dataset was allocated for training, while the remaining 20\% was reserved for validation to facilitate model selection. Once the best-performing model was identified during the validation phase, the entire training and development datasets were combined to retrain the final model. This final model was then evaluated on the test dataset, which was held out during the entire training process to ensure an unbiased assessment of the model's generalization performance. This approach adheres to standard practices in machine learning research to prevent data leakage and ensure robust evaluation \citep{Goodfellow-et-al-2016}.

@@

The model training process relied on Binary Cross-Entropy (BCE) as the loss function for Track A and C, and Cross-Entropy for Track B due to its suitability for classification tasks. Optimization was performed using the \texttt{AdamW} optimizer, which improves upon the standard Adam optimizer by decoupling weight decay and learning rate updates \citep{Loshchilov2019AdamW}. To further enhance training stability and convergence, a cosine annealing learning rate scheduler with restarts was employed. This scheduling approach helped adaptively reduce the learning rate over time, facilitating better exploration of the loss landscape and improving generalization. The model was trained for a fixed number of epochs, and early stopping was used to terminate training if the validation performance plateaued, thus avoiding overfitting.

>>>>

\subsection{Data Splits and Usage}
The dataset\cite{muhammad2025brighterbridginggaphumanannotated} was divided into three subsets: training, development (validation), and testing. Specifically, 80\% of the training dataset was allocated for training, while the remaining 20\% was reserved for validation to facilitate model selection. Once the best-performing model was identified during the validation phase, the entire training and development datasets were combined to retrain the final model. This final model was then evaluated on the test dataset, which was held out during the entire training process to ensure an unbiased assessment of the model's generalization performance. This approach adheres to standard practices in machine learning research to prevent data leakage and ensure robust evaluation \citep{Goodfellow-et-al-2016}.

@@

The model training process relied on Binary Cross-Entropy (BCE) as the loss function for Track A and C, and Cross-Entropy for Track B due to its suitability for classification tasks. Given the unbalanced dataset, a weighted loss approach was employed to ensure that the model adequately learned from all classes. Optimization was performed using the \texttt{AdamW} optimizer, which improves upon the standard Adam optimizer by decoupling weight decay and learning rate updates \citep{Loshchilov2019AdamW}. To further enhance training stability and convergence, a cosine annealing learning rate scheduler with restarts was employed. This scheduling approach helped adaptively reduce the learning rate over time, facilitating better exploration of the loss landscape and improving generalization. The model was trained for a fixed number of epochs, and early stopping was used to terminate training if the validation performance plateaued, thus avoiding overfitting.

### 2025-02-22 paper/sections/RelatedWork.tex
add conclusion

Moreover, datasets specifically curated for emotion detection in underrepresented languages have been explored. The datasets presented in \cite{muhammad2025brighterbridginggaphumanannotated} and \cite{belay-etal-2025-evaluatin} serve as essential resources for training models and evaluating performance in real-world settings. These datasets enable the training of robust models capable of handling linguistic diversity.

>>>>

Moreover, datasets specifically curated for emotion detection in underrepresented languages have been explored. The datasets presented in \cite{muhammad2025brighterbridginggaphumanannotated} and \cite{belay-etal-2025-evaluating} serve as essential resources for training models and evaluating performance in real-world settings. These datasets enable the training of robust models capable of handling linguistic diversity.

### 2025-02-22 paper/sections/SystemOverview.tex
add conclusion

Following feature extraction, we applied multiple classification algorithms to perform the multi-label classification task. One of the classifiers used was the Multi-Layer Perceptron (MLP), a feedforward artificial neural network capable of modeling complex relationships between the extracted features and the target labels. Additionally, the system employed XGBoost, a gradient boosting framework renowned for its effectiveness in structured data classification \citep{chen2016xgboost}. Furthermore, Support Vector Machines (SVMs) were utilized as a classification method due to their ability to operate effectively in high-dimensional feature spaces by identifying optimal hyperplanes for classification \citep{cortes1995support}.

For languages with sufficient pretrained models available, we identified the best-performing embedding model and paired it with SVM as the classifier. This approach leverages the strengths of the pretrained embedding models in capturing language-specific nuances, while the SVM classifier ensures robust performance for multi-label classification. On the other hand, for languages with limited pretrained resources, we utilized the multilingual embedding model ``Multilingual-E5-large-instruct'' \citep{wang2024multilingual} in combination with XGBoost as the classifier. The ``Multilingual-E5-large-instruct'' model, designed to generalize across diverse languages, enabled the system to maintain high performance even in resource-constrained settings.

>>>>

Following feature extraction, we applied multiple classification algorithms to perform the multi-label classification task. One of the classifiers used was the Multi-Layer Perceptron (MLP), a feedforward artificial neural network capable of modeling complex relationships between the extracted features and the target labels. Additionally, the system employed XGBoost, a gradient boosting framework renowned for its effectiveness in structured data classification \citep{chen2016xgboost}. Furthermore, SVMs\footnote{Support Vector Machines} were utilized as a classification method due to their ability to operate effectively in high-dimensional feature spaces by identifying optimal hyperplanes for classification \citep{cortes1995support}.

For languages with sufficient pretrained models available using MTEB\footnote{Massive Text Embedding Benchmark}\citep{muennighoff2022mteb}, we identified the best-performing embedding model and paired it with SVM as the classifier. This approach leverages the strengths of the pretrained embedding models in capturing language-specific nuances, while the SVM classifier ensures robust performance for multi-label classification. On the other hand, for languages with limited pretrained resources, we utilized the multilingual embedding model "Multilingual E5 large instruct" \citep{wang2024multilingual} in combination with XGBoost as the classifier. The model, designed to generalize across diverse languages, enabled the system to maintain high performance even in resource-constrained settings.

### 2025-02-22 paper/tables/MainResults.tex
add conclusion

@@

    \caption{Results across Track A, B, and C showing macro-average F1 scores of Our Model , Paraticipants Best Model scores, Task Dataset Model\citep{muhammad2025brighterbridginggaphumanannotated} and rankings.}

@@

        \begin{tabular}{l|ccc|cccc|cccc}

@@

                              & \multicolumn{3}{c|}{Track A} & \multicolumn{4}{c|}{Track B} & \multicolumn{4}{c}{Track C}                                                                                                                                                     \\
            \cline{2-12}
            \textbf{Language} & \textbf{Our Model}               & \textbf{Best Model}          & \textbf{Rank}               & \textbf{Our Model} & \textbf{Best Model} & \textbf{Task Dataset Model} & \textbf{Rank} & \textbf{Our Model} & \textbf{Best Model} & \textbf{Task Dataset Model} & \textbf{Rank} \\

@@

            afr               & 54.01                        & 69.86                        & 15/37                       & —              & —                   & —                & —             & 54.01          & 70.50               & 61.28            & 4/14          \\
            amh               & 61.20                        & 77.31                        & 19/43                       & 49.42          & 85.58               & —                & 14/23         & 61.20          & 66.68               & —                & 6/13          \\
            arq               & 51.07                        & 66.87                        & 24/43                       & 36.54          & 63.38               & 36.37            & 18/26         & 51.07          & 65.35               & 55.75            & 6/14          \\
            ary               & 51.88                        & 62.92                        & 20/41                       & —              & —                   & —                & —             & 51.88          & 63.22               & 52.76            & 6/12          \\
            chn               & 56.65                        & 70.94                        & 30/41                       & 48.47          & 72.24               & 51.86            & 19/27         & 56.65          & 56.65               & 55.23            & 7/15          \\
            deu               & 60.60                        & 73.99                        & 25/50                       & 54.10          & 76.57               & 56.21            & 18/27         & 60.60          & 73.62               & 59.17            & 6/15          \\
            eng               & 73.97                        & 82.30                        & 35/96                       & 68.81          & 84.04               & 64.15            & 28/43         & 73.97          & 82.07               & 65.58            & 4/17          \\
            esp               & 76.19                        & 84.88                        & 26/47                       & 66.70          & 80.80               & 72.59            & 23/29         & 76.19          & 85.00               & 73.29            & 5/15          \\
            hau               & 63.22                        & 75.07                        & 18/40                       & 58.42          & 77.00               & 39.16            & 15/26         & 63.22          & 73.14               & 51.91            & 4/13          \\
            hin               & 80.32                        & 92.57                        & 33/43                       & —              & —                   & —                & —             & 80.32          & 92.16               & 79.73            & 6/16          \\
            ibo               & 50.93                        & 60.01                        & 14/34                       & —              & —                   & —                & —             & 50.93          & 60.47               & 37.40            & 4/12          \\
            ind               & —                            & —                            & —                           & —              & —                   & —                & —             & 35.64          & 67.24               & 57.29            & 15/17         \\
            jav               & —                            & —                            & —                           & —              & —                   & —                & —             & 25.62          & 25.62               & 50.47            & 12/13         \\
            kin               & 51.94                        & 65.74                        & 6/31                        & —              & —                   & —                & —             & 51.94          & 64.59               & 34.36            & 2/11          \\
            mar               & 81.10                        & 90.58                        & 23/42                       & —              & —                   & —                & —             & 81.10          & 90.42               & 77.24            & 6/13          \\
            orm               & 54.31                        & 61.64                        & 11/36                       & —              & —                   & —                & —             & 54.31          & 60.07               & —                & 3/11          \\
            pcm               & 53.09                        & 67.40                        & 22/34                       & —              & —                   & —                & —             & 53.09          & 67.40               & 48.67            & 5/10          \\
            ptbr              & 47.99                        & 68.33                        & 27/42                       & 38.20          & 71.00               & 46.72            & 22/25         & 47.99          & 68.36               & 51.60            & 8/14          \\
            ptmz              & 50.08                        & 54.77                        & 6/36                        & —              & —                   & —                & —             & 50.08          & 55.54               & 40.44            & 3/14          \\
            ron               & 73.75                        & 79.43                        & 14/43                       & 57.61          & 72.60               & 57.69            & 19/26         & 73.75          & 77.27               & 76.23            & 4/15          \\
            rus               & 82.42                        & 90.87                        & 33/50                       & 78.41          & 92.54               & 87.66            & 22/30         & 82.42          & 90.62               & 76.97            & 6/17          \\
            som               & 48.26                        & 57.65                        & 10/34                       & —              & —                   & —                & —             & 48.26          & 56.66               & —                & 4/13          \\
            sun               & 42.48                        & 54.97                        & 19/37                       & —              & —                   & —                & —             & 42.48          & 50.72               & 46.33            & 4/11          \\
            swa               & 29.52                        & 41.47                        & 15/32                       & —              & —                   & —                & —             & 29.52          & 38.43               & 33.27            & 5/14          \\
            swe               & 56.51                        & 62.62                        & 13/40                       & —              & —                   & —                & —             & 56.51          & 64.53               & 51.18            & 6/13          \\
            tat               & 64.32                        & 84.59                        & 18/36                       & —              & —                   & —                & —             & 64.32          & 83.59               & 60.66            & 5/11          \\
            tir               & 52.37                        & 59.05                        & 6/34                        & —              & —                   & —                & —             & 52.37          & 55.24               & —                & 2/10          \\
            ukr               & 48.62                        & 72.56                        & 29/40                       & 42.55          & 70.75               & 43.54            & 17/24         & 48.62          & 71.99               & 54.76            & 10/17         \\
            vmw               & 16.81                        & 32.50                        & 5/31                        & —              & —                   & —                & —             & 16.80          & 26.02               & 20.41            & 6/9           \\
            xho               & —                            & —                            & —                           & —              & —                   & —                & —             & 16.64          & 44.26               & 30.79            & 6/10          \\
            yor               & 34.09                        & 46.13                        & 9/32                        & —              & —                   & —                & —             & 34.09          & 46.79               & 27.44            & 4/11          \\
            zul               & —                            & —                            & —                           & —              & —                   & —                & —             & 16.35          & 16.35               & 22.03            & 8/11          \\

>>>>

% Multilingual-E5-large-instruct \citep{wang2024multilingual}
% AraBert Large Algerian \citep{arabert}

@@

    \caption{Results across Track A, B, and C showing macro-average F1 scores of Our Model , Paraticipants Best Model scores, Task Dataset Model(Baseline)\citep{muhammad2025brighterbridginggaphumanannotated} and rankings.}

@@

        \begin{tabular}{l|c|ccc|cccc|cccc}

@@

                                                    &                                                                                              & \multicolumn{3}{c|}{Track A} & \multicolumn{4}{c|}{Track B} & \multicolumn{4}{c}{Track C}                                                                                                                                         \\
            \cline{3-13}
            \textbf{Language}                       & \textbf{Our Model Name}                                                                      & \textbf{Ours}                & \textbf{Best}                & \textbf{Rank}               & \textbf{Ours} & \textbf{Best} & \textbf{Baseline} & \textbf{Rank} & \textbf{Ours} & \textbf{Best} & \textbf{Baseline} & \textbf{Rank} \\

@@

            Afrikaans(afr)                          & \citep{wang2024multilingual} + SVM                                                           & 54.01                        & 69.86                        & 15/37                       & —             & —             & —                 & —             & 54.01         & 70.50         & 61.28             & 4/14          \\
            Amharic(amh)                            & \citep{arabert}               + SVM                                                          & 61.20                        & 77.31                        & 19/43                       & 49.42         & 85.58         & —                 & 14/23         & 61.20         & 66.68         & —                 & 6/13          \\
            Algerian Arabic(arq)                    & \citep{wang2024multilingual}   + SVM                                                         & 51.07                        & 66.87                        & 24/43                       & 36.54         & 63.38         & 36.37             & 18/26         & 51.07         & 65.35         & 55.75             & 6/14          \\
            Moroccan Arabic(ary)                    & \citep{wang2024multilingual}   + SVM                                                         & 51.88                        & 62.92                        & 20/41                       & —             & —             & —                 & —             & 51.88         & 63.22         & 52.76             & 6/12          \\
            Chinese(chn)                            & ZPoint Large\footnote{\url{https://huggingface.co/iampanda/zpoint_large_embedding_zh}} + SVM & 56.65                        & 70.94                        & 30/41                       & 48.47         & 72.24         & 51.86             & 19/27         & 56.65         & 56.65         & 55.23             & 7/15          \\
            German(deu)                             & \citep{wang2024multilingual}       + SVM                                                     & 60.60                        & 73.99                        & 25/50                       & 54.10         & 76.57         & 56.21             & 18/27         & 60.60         & 73.62         & 59.17             & 6/15          \\
            English(eng)                            & \citep{zhang2025jasperstelladistillationsota}                                                & 73.97                        & 82.30                        & 35/96                       & 68.81         & 84.04         & 64.15             & 28/43         & 73.97         & 82.07         & 65.58             & 4/17          \\
            Spanish(esp)                            & \citep{wang2024multilingual}      + SVM                                                      & 76.19                        & 84.88                        & 26/47                       & 66.70         & 80.80         & 72.59             & 23/29         & 76.19         & 85.00         & 73.29             & 5/15          \\
            Hausa(hau)                              & \citep{dobler-demelo-2023-focus} + SVM                                                       & 63.22                        & 75.07                        & 18/40                       & 58.42         & 77.00         & 39.16             & 15/26         & 63.22         & 73.14         & 51.91             & 4/13          \\
            Hindi(hin)                              & \citep{wang2024multilingual}  + SVM                                                          & 80.32                        & 92.57                        & 33/43                       & —             & —             & —                 & —             & 80.32         & 92.16         & 79.73             & 6/16          \\
            Igbo(ibo)                               & \citep{wang2024multilingual}   + SVM                                                         & 50.93                        & 60.01                        & 14/34                       & —             & —             & —                 & —             & 50.93         & 60.47         & 37.40             & 4/12          \\
            Indonesian(ind)                         & \citep{wang2024multilingual}   + XGB                                                         & —                            & —                            & —                           & —             & —             & —                 & —             & 35.64         & 67.24         & 57.29             & 15/17         \\
            Javanese(jav)                           & \citep{wang2024multilingual}   + XGB                                                         & —                            & —                            & —                           & —             & —             & —                 & —             & 25.62         & 25.62         & 50.47             & 12/13         \\
            Kinyarwanda(kin)                        & \citep{wang2024multilingual}  + SVM                                                          & 51.94                        & 65.74                        & 6/31                        & —             & —             & —                 & —             & 51.94         & 64.59         & 34.36             & 2/11          \\
            Marathi(mar)                            & \citep{wang2024multilingual}   + SVM                                                         & 81.10                        & 90.58                        & 23/42                       & —             & —             & —                 & —             & 81.10         & 90.42         & 77.24             & 6/13          \\
            Oromo(orm)                              & \citep{wang2024multilingual}    + SVM                                                        & 54.31                        & 61.64                        & 11/36                       & —             & —             & —                 & —             & 54.31         & 60.07         & —                 & 3/11          \\
            Nigerian-Pidgin(pcm)                    & \citep{wang2024multilingual}   + SVM                                                         & 53.09                        & 67.40                        & 22/34                       & —             & —             & —                 & —             & 53.09         & 67.40         & 48.67             & 5/10          \\
            Pt\footnote{Portuguese} Brazilian(ptbr) & \citep{souza2020bertimbau} + SVM                                                             & 47.99                        & 68.33                        & 27/42                       & 38.20         & 71.00         & 46.72             & 22/25         & 47.99         & 68.36         & 51.60             & 8/14          \\
            Pt Mozambican(ptmz)                     & \citep{wang2024multilingual}   + SVM                                                         & 50.08                        & 54.77                        & 6/36                        & —             & —             & —                 & —             & 50.08         & 55.54         & 40.44             & 3/14          \\
            Romanian(ron)                           & \citep{wang2024multilingual}     + SVM                                                       & 73.75                        & 79.43                        & 14/43                       & 57.61         & 72.60         & 57.69             & 19/26         & 73.75         & 77.27         & 76.23             & 4/15          \\
            Russian(rus)                            & \citep{snegirev2025russianfocusedembeddersexplorationrumteb}     + SVM                       & 82.42                        & 90.87                        & 33/50                       & 78.41         & 92.54         & 87.66             & 22/30         & 82.42         & 90.62         & 76.97             & 6/17          \\
            Somali(som)                             & \citep{wang2024multilingual}       + SVM                                                     & 48.26                        & 57.65                        & 10/34                       & —             & —             & —                 & —             & 48.26         & 56.66         & —                 & 4/13          \\
            Sundanese(sun)                          & \citep{wang2024multilingual}      + SVM                                                      & 42.48                        & 54.97                        & 19/37                       & —             & —             & —                 & —             & 42.48         & 50.72         & 46.33             & 4/11          \\
            Swahili(swa)                            & \citep{wang2024multilingual}      + SVM                                                      & 29.52                        & 41.47                        & 15/32                       & —             & —             & —                 & —             & 29.52         & 38.43         & 33.27             & 5/14          \\
            Swedish(swe)                            & \citep{wang2024multilingual}     + SVM                                                       & 56.51                        & 62.62                        & 13/40                       & —             & —             & —                 & —             & 56.51         & 64.53         & 51.18             & 6/13          \\
            Tatar(tat)                              & \citep{wang2024multilingual}    + SVM                                                        & 64.32                        & 84.59                        & 18/36                       & —             & —             & —                 & —             & 64.32         & 83.59         & 60.66             & 5/11          \\
            Tigrinya(tir)                           & \citep{wang2024multilingual}    + SVM                                                        & 52.37                        & 59.05                        & 6/34                        & —             & —             & —                 & —             & 52.37         & 55.24         & —                 & 2/10          \\
            Ukrainian(ukr)                          & \citep{sturua2024jinaembeddingsv3multilingualembeddingstask}   + SVM                         & 48.62                        & 72.56                        & 29/40                       & 42.55         & 70.75         & 43.54             & 17/24         & 48.62         & 71.99         & 54.76             & 10/17         \\
            Emakhuwa(vmw)                           & \citep{sturua2024jinaembeddingsv3multilingualembeddingstask}   + SVM                         & 16.81                        & 32.50                        & 5/31                        & —             & —             & —                 & —             & 16.80         & 26.02         & 20.41             & 6/9           \\
            isiXhosa(xho)                           & \citep{wang2024multilingual}     + XGB                                                       & —                            & —                            & —                           & —             & —             & —                 & —             & 16.64         & 44.26         & 30.79             & 6/10          \\
            Yoruba(yor)                             & \citep{wang2024multilingual}     + SVM                                                       & 34.09                        & 46.13                        & 9/32                        & —             & —             & —                 & —             & 34.09         & 46.79         & 27.44             & 4/11          \\
            isiZulu(zul)                            & \citep{wang2024multilingual}     + XGB                                                       & —                            & —                            & —                           & —             & —             & —                 & —             & 16.35         & 16.35         & 22.03             & 8/11          \\
